---
title: "Modi"
author: "AlexSaltman"
date: "`r Sys.Date()`"
# runtime: shiny
output:
  html_document:
    code_folding: hide
    highlight: zenburn # pygments, kate, monochrome, zenburn
    toc: yes
params:
  use_paramfile:
    value: TRUE
    description: "Boolean value whether or not to load param file if present"
  paramfile:
    value: "../params/Modi.yaml"
  ROOTPATH:
    value: "/mnt/e/MSPC001012/"
  PROF_PATH:
    value: "./PROF/52603/52603_1_7"
    description: "directory location for profiling gct files"
  MODI_PATH:
    value: "./diGG/52604_2_7/"
    description: "directory location for modi gct files"
  moditype:
    value:
      gg
  amino_acids:
    value:
      [K]
  cluster_methods:
    value: [kmeans]
  formula:
    value: "~0+Extract_Genotype:Extract_Treatment"
    description: "will move to separate volcano.Rmd document in the future, probably"
  pca:
    value:
      color:
        value: Extract_Genotype
      marker:
        value: Extract_Treatment
      annotate:
        value: true
      center:
        value: true
      scale:
        value:
          true
  cluster:
    value:
      cluster_rows:
        value: true
        description: "cluster the rows"
      cluster_columns:
        value: true
      cluster_row_slices:
        value: true
      cluster_col_slices:
        value: true
      clustering_distance_rows:
        value: euclidean
      clustering_method_rows:
        value: ward.d2
      legend_include:
        value:
        description: "metadata variables to include"
      legend_exclude:
        value:
        description: "metadata variables to exclude"
  logging:
    value:
      message: TRUE
      warning: TRUE
      quiet: TRUE
end:
#

---


```{r label=modi-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(cmapR))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(limma))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(tidymodels))
library(tidymodels)
library(tibble)
library(magrittr)
source("funcs.R")

mykable <- function(data) {
    knitr::kable(data, booktabs = TRUE, digits = 4) %>%
        kable_styling(latex_options = c("striped", "scale_down"),
                      bootstrap_options = c("striped", 'condensed', "bordered")
                      )
}

get_param <- function(param){
  if(is.list(param) && "value" %in% names(param)){
    return(param$value)
  } else{
    return(param)
  }
}
```

# Load

## params

```{r, label=modi.load.params, class.source='fold-show'}

#.f <- params$paramfile
.f <- get_param(params$paramfile)
.use_paramfile <- get_param(params$use_paramfile)
if (!is.null(.f) && fs::file_exists(.f) && params$use_paramfile == TRUE) {
  subparams <- yaml::yaml.load_file(.f)
  cat(paste("loading", .f))
  #params <- modifyList(params, local_params)
} else{
  subparams <- params
}
# cat(paste(subparams))
# (params)
# (subparams)
```


## gct

```{r, label="modi.load.data", class.source='fold-show'}
# ROOTPATH <- fs::path_abs(subparams$ROOTPATH)
# setwd(ROOTPATH)

ROOTPATH <- get_param(subparams$ROOTPATH)
PROF_PATH <- fs::path(ROOTPATH, get_param(subparams$PROF_PATH))
MODI_PATH <- fs::path(ROOTPATH, get_param(subparams$MODI_PATH))
# modi_PATH <- fs::path_abs(subparams$modi_PATH)

#prof_files <- PROF_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")

prof_files <- PROF_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")
(prof_files)

modi_files <- MODI_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")
(modi_files)

# parse_gct_dropzeros <- function (file){
#   gct <- cmapR::parse_gctx(file)
#   mat <- gct@mat
#   mat_nonzero <- mat[apply(mat, 1, sum) != 0, ]
#   remaining_rids <- rownames(mat_nonzero)
#   gct <- gct %>% subset_gct(rid=remaining_rids)
#   if ("TaxonID" %in% colnames(gct@cdesc)){
#     gct@cdesc$TaxonID <- NULL
#   }
#   gct
#   # this has to go somewhere else and be more costomizable
#   # valid_cids <- gct@cdesc %>% filter(Extract_Genotype != "RefMix") %>% pull(id)
#   # gct <- gct %>% subset_gct(cid = valid_cids)
# }
# prof_gcts <- prof_files %>% purrr::map(parse_gct_dropzeros)

#prof_gcts <- prof_files %>% purrr::map(cmapR::parse_gctx)
prof_gcts <- prof_files %>% purrr::map(cmapR::parse_gctx)
modi_gcts <- modi_files %>% purrr::map(cmapR::parse_gctx)
# need to remove sample identifying info that is in the rdesc
cols_to_drop <- c("recno", "runno", "searchno", "basename", "TopScore", "TaxonID")
modi_gcts <- purrr::map(modi_gcts, function(gct) {
  gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
  return(gct)
})



modi_gcts <- purrr::map(modi_gcts, function(gct) {
  if ("taxon" %in% colnames(gct@rdesc) & (!"TaxonID" %in% colnames(gct@rdesc))) {
    gct@rdesc$TaxonID <- gct@rdesc$taxon 
  }
  return(gct)  # ensure the modified gct is returned
})

  
cdesc_cols_to_drop <- c("TaxonID")
modi_gcts <- purrr::map(modi_gcts, function(gct) {
  gct@cdesc <- gct@cdesc[, !colnames(gct@cdesc) %in% cdesc_cols_to_drop]
  return(gct)
})



cdesc_cols_to_drop <- c("TaxonID")
prof_gcts <- purrr::map(prof_gcts, function(gct) {
  gct@cdesc <- gct@cdesc[, !colnames(gct@cdesc) %in% cdesc_cols_to_drop]
  return(gct)
})
# need to remove sample identifying info that is in the rdesc
cols_to_drop <- c("recno", "runno", "searchno", "basename", "GPGroup", "PSMs", "PeptidePrint", "EXPRecNo", "EXPRunNo", "EXPSearchNo", "LabelFLAG", "PeptideCount")
prof_gcts <- purrr::map(prof_gcts, function(gct) {
  gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
  return(gct)
})


moditype <- get_param(subparams$moditype)
amino_acids <- get_param(subparams$amino_acids)

subset_modi <- function(modi_gct){
  .tokeep <- modi_gct@rdesc %>%
    dplyr::filter(AA %in% amino_acids & ModiType == moditype)  %>%
    rownames
  modi_gct_f <- cmapR::subset_gct(modi_gct, rid = .tokeep)
  return(modi_gct_f)
}
modi_gcts_filtered <- modi_gcts %>%
  purrr::map(subset_modi)


#gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
```


## normalize here

```{r label=normalize}


normalize <- function(gct){
  
  dfm <- gct %>% melt_gct()
  smallval <- min(dfm[ dfm$value != 0, "value" ], na.rm=T) / 2
  dfm[ dfm$value == 0, "value"] <- smallval
  dfm[ is.na(dfm$value), "value"] <- smallval
  #mutate(value = value / min(value[value != 0], na.rm=T)) %>% 
  dfm <- dfm %>% 
    group_by(id.y, TaxonID) %>%
    mutate(mednorm = (value / median(value, na.rm=TRUE) ) ) %>%
    mutate(logmednorm = log(mednorm)) %>%
    ungroup() %>% 
    mutate(logmednorm = logmednorm + abs(min(logmednorm)))
  mat_log <- dfm %>% pivot_wider(id_cols=id.x, names_from=id.y, values_from=logmednorm ) %>% as.data.frame
  rownames(mat_log) <- mat_log$id.x
  mat_log$id.x <- NULL
  gct_norm_log <- gct %>% cmapR::GCT(mat = as.matrix(mat_log), rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
  gct_norm_log
}

modi_gcts_norm_log <- modi_gcts_filtered %>% purrr::map(normalize)

prof_gcts_norm_log <- prof_gcts %>% purrr::map(normalize)


```

```{r, label='modi-dist'}
modi_gcts_norm_log[[1]] %>% melt_gct() %>% filter(TaxonID!="nan") %>% ggpubr::ggviolin(x='id.y', y='value', color = "TaxonID")
```

```{r, label='prof-dist'}
prof_gcts_norm_log[[1]] %>% melt_gct() %>% filter(!is.na(TaxonID) & TaxonID != "NaN") %>% ggpubr::ggviolin(x='id.y', y='value', color = "TaxonID")
```

```{r, label='modi-reduce-merge'}

modi_gct_merged <- base::Reduce(accumulate=F,
                    f=function(...) cmapR::merge_gct(..., dim="col"),
                    x=modi_gcts_norm_log
)

rownames(modi_gct_merged@cdesc) <- modi_gct_merged@cdesc$id # always called ID when melting, taken from rownames
rownames(modi_gct_merged@rdesc) <- modi_gct_merged@rdesc$id # always called ID when melting, taken from rownames


cidcounts <- modi_gct_merged %>% melt_gct %>% group_by(recno, runno, searchno) %>% summarise(count=n_distinct(id.y)) %>% ungroup
ridcounts <- modi_gct_merged %>% melt_gct %>% group_by(id.x) %>% summarise(count=n_distinct(id.y)) %>% ungroup
.table <- modi_gct_merged@rdesc %>% rownames_to_column("id.x")
ridcounts <- left_join(ridcounts, .table, by = 'id.x')
.maxval <- max(ridcounts$count)


# here filter for species
#.species_filter <- modi_gct_merged@rdesc   #no filter
.species_filter <- modi_gct_merged@rdesc %>% dplyr::filter(TaxonID=="10090")
ridcounts <- ridcounts %>% filter(GeneID %in% .species_filter$GeneID)

rids_filtered <- ridcounts %>% filter(count >= .maxval) %>% pull(id.x)
modi_gct <- subset_gct(modi_gct_merged, rid=rids_filtered)


prof_gct_merged <- base::Reduce(
  f = function(...) cmapR::merge_gct(..., dim = "col"),
  x=prof_gcts_norm_log
)
rownames(prof_gct_merged@cdesc) <- prof_gct_merged@cdesc$id # always called ID when melting, taken from rownames
rownames(prof_gct_merged@rdesc) <- prof_gct_merged@rdesc$id # always called ID when melting, taken from rownames
ridcounts <- prof_gct_merged %>% melt_gct %>% group_by(id.x) %>% summarise(count=n_distinct(id.y)) %>% ungroup
ridcounts$GeneID <- ridcounts$id.x #this is true for profiling data
.maxval <- max(ridcounts$count)

.species_filter <- prof_gct_merged@rdesc %>% dplyr::filter(TaxonID == "10090")
ridcounts <- ridcounts %>% filter(GeneID %in% .species_filter$GeneID)

rids_filtered <- ridcounts %>% filter(count >= .maxval) %>% pull(id.x)
#prof_gct <- prof_gct_merged
prof_gct <- subset_gct(prof_gct_merged, rid=rids_filtered)

# # here filter for species

```


## batch correction
```{r, label=modi.batch}

do_batch <- get_param(subparams$batchcorrection$do_batch_correction)
batchvar <- get_param(subparams$batchcorrection$batchvar)

if (do_batch == TRUE){
  
  batchvalues <- modi_gct@cdesc[[batchvar]] %>% as.character()
  mat_combat <- sva::ComBat(modi_gct@mat, batchvalues)
  modi_gct_batch <- cmapR::GCT(mat = mat_combat, cid = modi_gct@cid, cdesc = modi_gct@cdesc, rdesc = modi_gct@rdesc)
  modi_gct <- modi_gct_batch
  
  
  #batchinfo <- prof_gct@cdesc %>% mutate(id=factor(id, levels=colnames(mat_log), ordered=TRUE)) %>% arrange(id)
  #batchinfo <- prof_gct_log@cdesc[[batchvar]] %>% as.character()
  batchvalues <- prof_gct@cdesc[[batchvar]] %>% as.character()
  mat_combat <- sva::ComBat(prof_gct@mat, batchvalues)
  prof_gct_batch <- prof_gct %>% cmapR::GCT(mat = mat_combat, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
  #prof_gct_norm_log <- prof_gct_batch
  prof_gct <- prof_gct_batch

}
#else:
#modi_gct_norm_log <- modi_gct %>% cmapR::GCT(mat = mat_log, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
# if not true
```
```{r, label=modi-export-normalized}
# TODO calculate file name dynamically based on normtype and batch info
.file <- file.path(ROOTPATH, "MSPC000936_modi_log_batchcor")
modi_gct %>% write_gct(.file, precision = 8)

.file <- file.path(ROOTPATH, "MSPC000936_prof_log_batchcor")
prof_gct %>% write_gct(.file, precision = 8)

  #write_gct(.file, precision = 8)

#.file <- file.path(ROOTPATH, "MSPC000936_modi_log_batchcor.tsv")

```

```{r, label=modi.zscore}

zscore_gct <- function(gct){
  
  .tmp <- gct %>%
    melt_gct() %>%
    group_by(id.x) %>%
    mutate(zscore = myzscore(value)) %>%
    ungroup()
  mat_zscore <- .tmp %>%  pivot_wider(id_cols=id.x, names_from=id.y, values_from=zscore) %>% as.data.frame
  rownames(mat_zscore) <- mat_zscore$id.x
  mat_zscore$id.x <- NULL
  mat_zscore <- as.matrix(mat_zscore)
  
  gct_zscore <- gct %>% cmapR::GCT(mat = mat_zscore, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
  
}
modi_gct_zscore <- zscore_gct(modi_gct)
prof_gct_zscore <- zscore_gct(prof_gct)
```



# Data overview

## distributions { .tabset }
### prof log
```{r, label='prof-dist-log2', eval=T}
prof_gct %>% melt_gct() %>%
  filter(TaxonID != "NaN") %>% 
  ggviolin(x="id.y", y="value", add = "median_mad", color = "TaxonID")  %>%
  ggpar(x.text.angle=90, title = "prof gct log")

```
### prof zscore
```{r, label='prof-dist-log', eval=T}
prof_gct_zscore %>% melt_gct() %>%
  filter(TaxonID != "NaN") %>% 
  ggviolin(x="id.y", y="value", add = "median_mad", color = "TaxonID")  %>%
  ggpar(x.text.angle=90, title = "prof gct log")

```
### prof zscore dist
```{r, label='prof-dist-zscore', eval=T}
prof_gct_zscore %>% melt_gct() %>%
  filter(TaxonID != "NaN") %>% 
  ggviolin(x="id.y", y="value", add = "median_mad", color = "TaxonID")  %>%
  ggpar(x.text.angle=90, title="prof gct zscore")

```

### prof zscore taxon
```{r}

prof_gct_zscore %>% melt_gct() %>%
  filter(TaxonID != "NaN") %>% 
  ggviolin(x="TaxonID", y="value", add = "median_mad")  %>%
  ggpar(x.text.angle=90, title="prof gct zscore")
```

## distributions - modi { .tabset }




### modi
```{r, label='modi-dist-log', cache=TRUE, eval=T}
modi_gct %>% melt_gct() %>%
  filter(TaxonID != "NaN" & TaxonID != "nan") %>% 
  ggviolin(x="id.y", y="value", add = "median_mad", color = "TaxonID")  %>%
  ggpar(x.text.angle=90, title="modi gct")
```

### modi zscore
```{r, label='modi-dist-zscore', cache=TRUE, eval=T}
modi_gct_zscore %>% melt_gct() %>%
  filter(TaxonID != "NaN" & TaxonID != "nan") %>% 
  ggviolin(x="id.y", y="value", add = "median_mad", color = "TaxonID")  %>%
  ggpar(x.text.angle=90, title="modi gct zscore")
```

### modi zscore taxon
```{r, label='modi-dist-zscore-taxon', cache=TRUE, eval=T}
modi_gct_zscore %>% melt_gct() %>%
  filter(TaxonID != "NaN" & TaxonID != "nan") %>% 
  ggviolin(x="TaxonID", y="value", add = "median_mad", color = "TaxonID")  %>%
  ggpar(x.text.angle=90, title="modi gct zscore")
```



## heatmap def
```{r, label='modi-heatmap-def', include=T, eval=T}
make_heatmap <- function(.gct, show_row_names=F, cut_by=NULL, pos_sel=NULL, cluster_rows=T, cluster_columns=T, width=9, height=12, ...){
  .more_params = list(...)
  .negsel <- c("sample", "plex", "recno",
               "runno", "searchno", "assay", "replicate", "name",
               "lu",
               "ProtocolNo",
               "Washes",
               "Date",
               "Experimenter",
               "Separation_1",
               "Separation_1Detail",
               "Separation_2",
               "Separation_2Detail",
               # "X1.1",
               # "X1Detail.1",
               # "X2.1",
               # "X2Detail.1",
               # "X1.2",
               # "X1Detail.2",
               # "X2.2",
               # "Separation_2Detail"
               "MS_Instrument",
               "Exp_Date",
               "Exp_Experimenter",
               "Exp_ExpType",
               "EXPRunNo",
               "ExpSearchNo"
               )
  .column_df <- .gct@cdesc %>% select(-any_of(.negsel))
  

  column_split <- NULL
  if (!is.null(cut_by)) {
    for (.col in cut_by) {
      .column_df[[.col]] <- factor(.column_df[[.col]], ordered = TRUE)
    }
    #.column_df <- .column_df %>% arrange(!!cut_by)
    column_split <- lapply(.column_df[cut_by], as.factor)
  } 

  recycle_colors <- function(type = "qual", n = 8) {
    # original_palette <- brewer_pal(palette = palette_name)(min(n, 8))
    original_palette <- brewer_pal(type = "qual", palette = 2)(min(n, 8))
    recycled_palette <- rep_len(original_palette, length.out = n)
    return(recycled_palette)
  }

  # note we do not have the extra columns anymore
  .nocolor <- c("id")
  if (!is.null(pos_sel)) .cols <- pos_sel
  else .cols <- colnames(.column_df)
  .color_list <- .column_df[, .cols] %>%
    select(-any_of(.nocolor)) %>% 
    as.list() %>%
    map(~ recycle_colors(n = length(unique(.))))

  # fix this maybe if better
  for (col in names(.color_list)) {
    # print(col)
    # names(.color_list[[col]]) <-  unique(.column_df[[col]])
    names(.color_list[[col]]) <- unique(.column_df[[col]])
  }


  .top_annotation <- columnAnnotation(
    df = .column_df[ , names(.color_list) ] %>% as.data.frame(),
    col = .color_list
    # col = anno_block(gp = gpar(fill = .color_list$metavar1))
  )

  #mat_zscore <- apply(mat(.gct) %>% head(50), MARGIN=1, myzscore)
  # mat_zscore <- apply(mat(.gct), MARGIN=1, myzscore)
  # dim(mat_zscore)
  # mat_zscore <- t(mat_zscore)  # transpose back to original structure
  # colnames(mat_zscore) <- colnames(mat(.gct))  # Save column names as row names before transpose
  # mat_zscore <- mat_zscore[, rownames(.column_df)]
  #mat_zscore = mat_zscore[, rownames(.column_df)]

  #mat_zscore <- mat_zscore[, rownames(.column_df)] # make sure order is the same
  mat <- mat(.gct)
  mat <- mat[, .gct@cdesc$id] # make sure same order
  mat <- mat[, .column_df$id] # make sure same order

  .h <- ComplexHeatmap::Heatmap(
    mat,
    top_annotation = .top_annotation,
    column_split = column_split,
    clustering_distance_rows = dist_no_na,
    clustering_method_rows = 'ward.D2',
    clustering_method_columns = 'ward.D2',
    clustering_distance_columns = dist_no_na,
    column_title_side = "top",
    column_title_rot = 0,
    column_names_gp = gpar(fontsize = 9),
    row_names_gp = gpar(fontsize=5),
    border = TRUE,
    use_raster = T,
    show_row_names = show_row_names,
    show_row_dend = T,
    cluster_rows = cluster_rows,
    cluster_columns = cluster_columns,
    width = width,
    height = height,
    ...,
  )
  .h
}
```


```{r, label="modi.modi_alldata_heatmap", eval=T, fig.height=8}
#cols_to_annotate <- c("celltissue", "genotype")
#cols_to_annotate <- c("Genotype", "recno")
cols_to_annotate <- c("Extract_Treatment", "Extract_Genotype")
# make_heatmap(modi_gct_batch, pos_sel = cols_to_annotate)
make_heatmap(modi_gct_zscore, pos_sel = cols_to_annotate, column_title = "modi_zscore")
```

```{r, label="modi-alldata-prof-heatmap", eval=T}
#cols_to_annotate <- c("celltissue", "genotype")
#cols_to_annotate <- c("Genotype", "recno")
cols_to_annotate <- c("Extract_Treatment", "Extract_Genotype")
# make_heatmap(modi_gct_batch, pos_sel = cols_to_annotate)
make_heatmap(prof_gct_zscore, pos_sel = cols_to_annotate, column_title = "prof_zscore")
```


```{r, label="modi.somemplot", eval=F}
prof_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```

```{r, label="modi.someotherplot", eval=F}
prof_correct_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```




# heatmaps


## heatmap prof
```{r, label='modi-heatmap-prof', cache=FALSE, eval=T}

pos_sel <- c("Extract_Treatment", "Extract_Genotype")
cut_by <- "Extract_Genotype"
.width = 8
.height = 14 
p <- make_heatmap(prof_gct, pos_sel = pos_sel, cut_by = NULL, show_row_names = F, column_title="log_norm_batch_corr")
print(p )
```

## heatmap modi
```{r, label='modi-heatmap-modi', cache=FALSE, eval=T}

pos_sel <- c("Extract_Treatment", "Extract_Genotype")
cut_by <- "Extract_Genotype"
p <- make_heatmap(modi_gct, pos_sel = pos_sel, cut_by = NULL, show_row_names = F,
                  column_title="log_norm_batch_corr"
                  )
print(p )
#make_heatmap(modi)
#make_heatmap(prof)
```



# PCA
## modi
render PCA.Rmd
```{r, label=modi-pca-render-func, message=T, warning=T, results='show', eval=T, class.source='fold-hide', eval=T}

#.params <- params$PCA

# .params$method <- "kmeans"

render_pca_report <- function(center = T, scale = F, title="PCA", color=NULL, marker=NULL, ...) {
  .output_file_name <- paste0("PCA", "title_", title, "center_", center, "scale_", scale, ".html")
  #.output_file_name <- fs::path(params$ROOTPATH, .output_file_name)
  
  # Capture additional arguments into a list
  additional_params <- list(...)

  rmarkdown::render("PCA.Rmd",
    output_file = .output_file_name,
    quiet = T,
    params = c(list(
      center = center,
      scale = scale,
      title = title,
      color = color,
      marker = marker,
      use_paramfile = F
      ),
      additional_params
    )
    # params=.params,
    # envir = globalenv(),
    # envir = new_env
    # documentation=1,
  )
  return(.output_file_name)
  # c(params$ROOTPATH, .output_file_name) %>% fs::path_join() %>% fs::path_abs()
}
```

```{r, label="modi.pca.modi.scaled", echo=T, eval=T}
# TODO: allow customization
gct <- modi_gct
#gct <- prof

.center <- T
.scale <- T
.pca_params <- get_param(subparams$pca)
# get subset pca params 
# there is no check here to see if it is defined in the yaml file, 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
.annotate <- .pca_params$annotate %>% get_param()
#.annotate <- T

.out <- render_pca_report(center = .center,
                          scale = .scale,
                          title="PCA_modipho",
                          color = .color,
                          marker = .marker,
                          annotate = .annotate
)
```
LINK
```{r, label="pca.report.1.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

```{r, label="modi.pca.modi.noscaled", echo=T, eval=T}
gct <- modi_gct

.center <- T
.scale <- F
.pca_params <- get_param(subparams$pca)
# get subset pca params 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
# .annotate <- .pca_params$annotate %>% get_param() # why is this n ot working?
.annotate <- T


.out <- render_pca_report(center = .center,
                          scale = .scale,
                          color = .color,
                          marker = .marker,
                          annotate = .annotate,
                          title="PCA_modipho"
                          )

# html
# .html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)

# htmltools::includeHTML(.out)
```

LINK
```{r, label="pca.report.2.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

## prof

```{r, label="modi.pca.prof.scaled", echo=T, eval=T}
# TODO: allow customization
gct <- prof_gct
#gct <- prof

.center <- T
.scale <- T
.pca_params <- get_param(subparams$pca)
# get subset pca params 
# there is no check here to see if it is defined in the yaml file, 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
.annotate <- .pca_params$annotate %>% get_param()
#.annotate <- T

.out <- render_pca_report(center = .center,
                          scale = .scale,
                          title="PCA_profiling",
                          color = .color,
                          marker = .marker,
                          annotate = .annotate
)
```
LINK
```{r, label="pca.report.3.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

```{r, label="modi.pca.prof.noscaled", echo=T, eval=T}
gct <- prof_gct

.center <- T
.scale <- F
.pca_params <- get_param(subparams$pca)
# get subset pca params 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
# .annotate <- .pca_params$annotate %>% get_param() # why is this n ot working?
.annotate <- T


.out <- render_pca_report(center = .center,
                          scale = .scale,
                          title = "PCA_profiling",
                          color = .color,
                          marker = .marker,
                          annotate = .annotate
                          )

# html
# .html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)

# htmltools::includeHTML(.out)
```

LINK
```{r, label="pca.report.4.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

## examine particular sites



```{r, label='modi.makeptable', eval=F}
make_p_table <- function(geneids){
  prof_gids <- intersect(geneids, prof_gct@rid)
  prof_expression_data <- mat(prof_gct)[prof_gids,]
  
  modi_gids <- intersect(geneids, modi_gct_norm_log@rdesc$GeneID)
  .psites <- modi_gct_norm_log@rdesc[modi_gct_norm_log@rdesc$GeneID %in% geneids, ]
  modi_expression_data <- mat(modi_gct_norm_log)[rownames(.psites), ]
  #modi_site <- modi@rdesc$modimodiites  # Adjust this line based on how modimodiites is stored in your data
  #prof$modimodiites <- modi_site
  DT::datatable(.psites)
}
```

```{r, label='modi-geneid-selection', eval=F}
  geneid <- "38418"
  make_p_table(geneid)

```




# Differential Analysis

## Limma
TODO: move to separate Rmd file

```{r, label='modi.limma.def', eval=T}

formula <- get_param(subparams$formula) %>% as.formula()
( contrasts_str <- get_param(subparams$contrasts) )
( contrasts_list <- strsplit(contrasts_str, "\n")[[1]] ) # strsplit returns vector array length 1
( contrasts_list <- purrr::map(contrasts_list, 
                               ~gsub("^['\"]|['\"]$", "", .) )
                               ) %>% unlist

do_limma <- function(gct){
#mod <- model.matrix(~0 + genotype, modi@cdesc)
  mod <- model.matrix(formula, gct@cdesc)
  colnames(mod) <- make.names(colnames(mod)) #where is this used
  limma_input <- gct %>% mat()
  limma_input <- limma_input[ ,rownames(mod) ]
  #contrast_list <- c(
  #  "genotypeIntS6 - genotypeWT",
  #  "genotypeIntS12 - genotypeWT",
  #  "genotypeIntS6 - genotypeIntS12",
  #  #"genotypeIntS6 + genotypeIntS12"
  #  )
  # contrasts <- makeContrasts(contrasts=contrast_list, levels = mod) #need to make variable contrast def
  contrasts <- makeContrasts(contrasts = contrasts_list, levels = mod)
  prefit <- lmFit(limma_input, mod)
  fit <- contrasts.fit(prefit, contrasts) %>% eBayes(trend = T, robust = T)
  fit
}
```


```{r, label='modi-limma-modi-run', eval=T}
modi_gct_norm_log <- modi_gct

fit <- do_limma(modi_gct_norm_log)
limma_res_modi <- purrr::map_dfr(
  contrasts_list,
  ~ topTable(fit, n=Inf, sort.by="none", coef=.x, confint = F  ) %>%
  rownames_to_column("GeneSymbolSite") %>%
  mutate(contrast=.x)
)
 
limma_res_modi_withmeta <- limma_res_modi %>% left_join(modi_gct@rdesc,
                             by = c('GeneSymbolSite' = 'id')
                             )
.order <- c("SiteName",
            "SiteID",
            "GeneID",
            "GeneSymbol",
            "GeneDescription",
            "Site",
            "AA",
            "MassShift",
            "Protein_Accession_Selected",
            "Protein_Length_Selected",
            "Reason_Selected",
            "AApos_Selected",
            "Protein_Accessions",
            "AApos_List",
            "TopScore"
            )
.front <- intersect(.order, colnames(limma_res_modi_withmeta))

# Reorder the dataframe
limma_res_modi_withmeta <- limma_res_modi_withmeta %>%
  select(all_of(.front), everything())

#.order <- c(
#  
#)
.file <- file.path(ROOTPATH, "MSPC000936_limma_res_modi.tsv")
limma_res_modi_withmeta %>% write_tsv(.file)

limma_res_modi_wide <- limma_res_modi %>%  pivot_wider(id_cols = GeneSymbolSite, names_from = contrast,
                               values_from = c("logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
                               )
.file <- file.path(ROOTPATH, "MSPC000936_limma_res_modi_wide.tsv")
limma_res_modi_wide %>% write_tsv(.file)

```

```{r, label='modi-limma-prof-run', eval=T}
fit <- do_limma(prof_gct)
limma_res_prof <- purrr::map_dfr(
  contrasts_list,
  ~ topTable(fit, n=Inf, sort.by="none", coef=.x, confint = F  ) %>%
  rownames_to_column("GeneID") %>%
  mutate(contrast=.x)
)
limma_res_prof_withmeta <- limma_res_prof %>% left_join(prof_gct@rdesc,
                             by = c('GeneID' = 'id')
                             )
.file <- file.path(ROOTPATH, "MSPC000936_limma_res_prof.tsv")
limma_res_prof_withmeta %>% write_tsv(.file)


limma_res_prof_wide <- limma_res_prof_withmeta %>% pivot_wider(id_cols = GeneID, names_from = contrast,
                               values_from = c("logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
                               )
.file <- file.path(ROOTPATH, "MSPC000936_limma_res_prof_wide.tsv")
limma_res_prof_wide %>% write_tsv(.file)
```

```{r, label='modi.prof.limma.merge', eval=T}
modi_prof_stats <- left_join(limma_res_modi_withmeta,
          limma_res_prof_withmeta,
          by = c("GeneID", "contrast"),
          suffix = c(".modi", ".prof")
          ) %>% 
  mutate(
    adj.P.Val.modi = replace_na(adj.P.Val.modi, 1),
    adj.P.Val.prof = replace_na(adj.P.Val.prof, 1),
    logFC.modi = replace_na(logFC.modi, 0),
    logFC.prof = replace_na(logFC.prof, 0),
  )

modi_prof_stats %<>% mutate(
  SigCat = if_else(  !is.na(adj.P.Val.modi) & adj.P.Val.modi < 0.05 & !is.na(adj.P.Val.prof) & adj.P.Val.prof < 0.05, "Both Sig",
                     if_else(
                       adj.P.Val.modi < 0.05 & adj.P.Val.prof >= 0.05, "modi Only",  
                       if_else(
                         adj.P.Val.modi >= 0.05 & adj.P.Val.prof < 0.05, "Prof Only",  
                         "Not Sig."
                     )
                     )
                     )
  
)
modi_prof_stats %>% count(SigCat)
```

## results { .tabset }

```{r, label='modi.2d.differential', eval=T, fig.width=5, fig.height=5, results='asis', cache=F}
#modi_prof_stats %>% 

plot_func <- function(df, contrast){
  cat(paste0("\n### ", contrast, "\n"))
  
  .df <- df %>% filter(contrast == contrast)
  
   p <- ggpubr::ggscatter(
     df,
    x="logFC.modi",
    y="logFC.prof",
    color = "SigCat",
    alpha=.25,
  )
   
  print(p)
  cat("\n")  # To ensure separation
}
#plot_func(modi_prof_stats, unique(modi_prof_stats$contrast))
unique( modi_prof_stats$contrast ) %>% walk( ~ plot_func( df = modi_prof_stats, contrast = . ) )

```


## top 50 { .tabset }

```{r, label='modi-top50', results='asis', fig.width=11, fig.height=16, cache=F}
contrast1 <- c(modi_prof_stats$contrast %>% unique)[1]

plot_contrast_top50 <- function(contrast){
  
  cat(paste0("\n### ", contrast, "\n"))
  #print(contrast)
  .sel <- modi_prof_stats %>% 
    filter(contrast == !!contrast) %>% 
    arrange(P.Value.modi) %>% 
    head(50)
  geneids <- .sel$GeneID
  .rids <- modi_gct_norm_log@rdesc %>% filter(GeneID %in% geneids) %>% pull(id)
  #make_p_table(geneids)
  .gct <-subset_gct(modi_gct_zscore, rid=.rids)
  
  pos_sel <- c("Extract_Treatment", "Extract_Genotype")
  #cut_by <- "Extract_Treatment"
  cut_by <- "Extract_Genotype"
  .width = 9
  .height = 16 
  p <- .gct %>% make_heatmap(pos_sel = pos_sel, cut_by = cut_by, show_row_names = T, column_title=contrast,
                             width = .width, height = .height
                             )
  print(p)
  p
  cat("\n")  # To ensure separation

}
modi_prof_stats$contrast %>% unique() %>% map(~plot_contrast_top50(.))
```




```{r, label=modi.geneset.heatmap, fig.width=15, fig.height=9}

.smarcs_h <- modi_gct_zscore@rdesc %>% filter(str_detect(GeneSymbol, "SMARC"))
.smarcs_m <- modi_gct_zscore@rdesc %>% filter(str_detect(GeneSymbol, "Smarc"))
.smarcs <- c(.smarcs_h$id,
             .smarcs_m$id
             )
pos_sel <- c("Extract_Treatment", "Extract_Genotype")
cut_by <- "Extract_Treatment"
cut_by <- "Extract_Genotype"
subset_gct(modi_gct_zscore, rid=.smarcs) %>% make_heatmap(show_row_names=T, pos_sel = pos_sel, cut_by = cut_by )
```

# params
```{r, label=modi.params}
( subparams )

```

