---
title: "PHOS"
author: "AlexSaltman"
date: "`r Sys.Date()`"
# runtime: shiny
output:
  html_document:
    code_folding: show
    highlight: zenburn # pygments, kate, monochrome, zenburn
    toc: yes

params:
  PROF:
    #value: "MSPC000967_TMT10_prof_Config_data_gct_linear_all_1more_nonzero_normtype_none_n10x7507.gct"
    value: "MSPC000967_TMT10_prof_Config_data_gct_all_1more_nonzero_normtype_median_n10x7507.gct"
    description: "gct file for one profiling experiment"
  PROF_PATH:
    value: "./exports/prof"

    description: "directory location for profiling gct files"
  PHOS:
    value: "51963_1_6_site_table_nr.tsv_n10x14785.gct"
    description: "gct file for one phosphoprofiling experiment"
  PHOS_PATH:
    value: "IMAC/"
    description: "directory location for phos gct files"

  ROOTPATH:
    #value: "`normalizePath(dirname('.'))`"
    value: "/media/alex/20230612/MSPC000936/MSPC000936/"
  paramfile:
    value: ""
  cluster_methods:
    value: [kmeans]
  pca:
    value:
      color:
        value:
      marker:
        value:
      annotate:
        value: true
      center:
        value: true
      scale:
        value:
          true
  cluster:
    value:
      cluster_rows:
        value: true
        description: "cluster the rows"
      cluster_columns:
        value: true
      cluster_row_slices:
        value: true
      cluster_col_slices:
        value: true
      clustering_distance_rows:
        value: euclidean
      clustering_method_rows:
        value: ward.d2
      legend_include:
        value:
        description: "metadata variables to include"
      legend_exclude:
        value:
        description: "metadata variables to exclude"
  logging:
    value:
      message: TRUE
      warning: TRUE
      quiet: TRUE
end:
#

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(cmapR))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(limma))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(tidymodels))
library(tidymodels)
library(tibble)
library(magrittr)
source("funcs.R")

mykable <- function(data) {
    knitr::kable(data, booktabs = TRUE, digits = 4) %>%
        kable_styling(latex_options = c("striped", "scale_down"),
                      bootstrap_options = c("striped", 'condensed', "bordered")
                      )
}
```


# Load

## gct

```{r, label="phos.load"}
ROOTPATH<- fs::path_abs(params$ROOTPATH)
PROF_PATH <- fs::path(ROOTPATH, params$PROF_PATH)
PHOS_PATH <- fs::path(ROOTPATH, params$PHOS_PATH)
PHOS_PATH <- fs::path_abs(params$PHOS_PATH)

prof_files <- PROF_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")

phos_files <- PHOS_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")
phos_gcts <- phos_files %>% purrr::map(cmapR::parse_gctx)
# need to remove sample identifying info that is in the rdesc
cols_to_drop <- c("recno", "runno", "searchno", "basename", "TopScore")
phos_gcts <- purrr::map(phos_gcts, function(gct) {
  gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
  return(gct)
})


#gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]

#phos_gcts %<>% purrr::map(~cmapR::subset_gct())
# filter out not STY aa modi
phos_gcts_f <- phos_gcts %>%  purrr::map(~cmapR::subset_gct(., rid = .@rdesc%>%dplyr::filter(AA %in% c("S", "T", "Y")) %>% rownames)  )
phos_gct_merged <- base::Reduce(accumulate=F,
                    f=function(...) cmapR::merge_gct(..., dim="col"),
                    x=phos_gcts_f
)

#phos_gcts %>% map(~.@cid)
#jphos_gcts %>% map(~ .@cdesc %>% head(2))

                    #x=phos_gcts[2:length(phos_gcts)],
                    #init = phos_gcts[[1]]

cidcounts <- phos_gct_merged %>% melt_gct %>% group_by(recno, runno, searchno) %>% summarise(count=n_distinct(id.y))
ridcounts <- phos_gct_merged %>% melt_gct %>% group_by(id.x) %>% summarise(count=n_distinct(id.y))
.maxval <- 30
.maxval <- max(ridcounts$count)
rids_filtered <- ridcounts %>% filter(count >= .maxval) %>% pull(id.x)

phos_gct <- subset_gct(phos_gct_merged, rid=rids_filtered)

```
## add metadata


```{r}
dfm <- phos_gct %>% melt_gct()
dfm %<>% group_by(id.y) %>%
  mutate(mednorm=value / median(value, na.rm=TRUE)) %>%
  mutate(logmednorm=log(mednorm+1)) %>%
  ungroup() %>%
  group_by(id.x) %>%
  mutate(zscore=scale(logmednorm, center=T, scale=F)) %>%
  ungroup

mat <- dfm %>% pivot_wider(id_cols=id.x, names_from=id.y, values_from=logmednorm ) %>% as.data.frame
rownames(mat) <- mat$id.x
mat$id.x <- NULL
#colnames(mat)

batchinfo <- phos_gct@cdesc %>% mutate(id=factor(id, levels=colnames(mat), ordered=TRUE)) %>% arrange(id)
batchvar <- batchinfo$recno %>% as.character()
mat_combat <- sva::ComBat(mat, batchvar)
phos_gct_batch <- phos_gct %>% cmapR::GCT(mat = mat_combat, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)


.tmp <- phos_gct_batch %>%
  melt_gct() %>%
  group_by(id.x) %>%
  mutate(zscore=myzscore(value)) %>%
  ungroup()
mat_zscore <- .tmp %>%  pivot_wider(id_cols=id.x, names_from=id.y, values_from=zscore) %>% as.data.frame
rownames(mat_zscore) <- mat_zscore$id.x
mat_zscore$id.x <- NULL
mat_zscore <- as.matrix(mat_zscore)

phos_gct_zscore <- phos_gct_batch %>% cmapR::GCT(mat = mat_zscore, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
```


# end July 19

```{r, label='phos.manual-metadata', eval=T}
.res<-left_join(prof@rdesc,
          distinct(phos@rdesc, GeneID, GeneSymbol),
          by = c('rdesc' = 'GeneID')
          )
prof@rdesc <- .res

.phos_metadata <- tibble::tribble(
~id, ~name,	~recno,	~runno,	~searchno,	~label,	~assay,	~genotype,	~replicate,
"51963_1_6_126", "DL1-1",	  51963,	1,	6,	"126", 	"IMAC",	"WT",	1,
"51963_1_6_127_N", "DL1-2",	  51963,	1,	6,	"127N",	"IMAC",	"WT",	2,
"51963_1_6_127_C", "DL1-3",	  51963,	1,	6,	"127C",	"IMAC",	"WT",	3,
"51963_1_6_128_N", "IntS6-1",	51963,	1,	6,	"128N",	"IMAC",	"IntS6",	1,
"51963_1_6_128_C", "IntS6-2",	51963,	1,	6,	"128C",	"IMAC",	"IntS6",	2,
"51963_1_6_129_N", "IntS6-3",	51963,	1,	6,	"129N",	"IMAC",	"IntS6",	3,
"51963_1_6_129_C", "IntS6-4",	51963,	1,	6,	"129C",	"IMAC", "IntS6", 	4,
"51963_1_6_130_N", "IntS12-1",	51963, 	1, 	6,	"130N",	"IMAC", "IntS12",	1,
"51963_1_6_130_C", "IntS12-2",	51963, 	1, 	6,	"130C",	"IMAC", "IntS12",	2,
"51963_1_6_131_N", "IntS12-3",	51963, 	1, 	6,	"131",	"IMAC", "IntS12",	3
)
#.phos_metadata$id <- .phos_metadata$name
.phos_sample_metadata <- .phos_metadata %>% mutate(id = factor(id, levels=phos@cid, ordered=T)) %>% arrange(id) %>% as.data.frame()
rownames(.phos_sample_metadata) <- .phos_sample_metadata$id
#phos@cdesc <- .phos_sample_metadata
#rownames(.phos_metadata) <- .phos_metadata$name #
#.cols <- c("id", "recno", "runno", "searchno", "label", "")
#out <- .phos_metadata %>% left_join(prof@cdesc, by = "id")
.the_order_we_want <- colnames(mat(phos))
.newnames <- .phos_sample_metadata %>%
   mutate(id=factor(id, levels(.the_order_we_want), ordered = T)) %>%
   arrange(id) %>%
   pull(name)
rownames(.phos_sample_metadata) <- .newnames
.phos_sample_metadata$id <- rownames(.phos_sample_metadata)
# source('funcs.R')
#colnames(mat(phos)) <- .newnames
#phos@cid <- .newnames
.new <- mat(phos)
colnames(.new) <- .newnames
# .new %>% dim
# phos %>% mat %>% dim
# phos@rdesc %>% dim
.rdesc <- phos@rdesc
.rid <- phos@rid

.phos <- new("GCT",
    mat=.new,
    cdesc=.phos_sample_metadata,
    cid=rownames(.phos_sample_metadata),
    #rid=phos@rid,
    #rid=rownames(.new),
    rid=.rid,
    rdesc=.rdesc,
    )
# subset_gct(phos)
# subset_gct(.phos)
phos <- .phos

#phos <- make_new_gct(phos, mat=.new, cdesc=.phos_metadata, cid=.phos_metadata$name)
#phos@cdesc <- .phos_metadata
# phos2 <- make_new_gct(phos, .new)
# phos <- phos2
```

```{r, label="phos-phos-debug"}

print(phos@cdesc)
print(phos@cid)
print(head(mat(phos)))
```


## preview metadata
 prof

```{r, label="phos-prof-mat"}
mat(prof) %>% head %>% mykable()
```

```{r, label="phos-prof-cdesc"}
prof@cdesc %>% mykable()
```


 phos

```{r, label="phos-phos-mat"}
mat(phos) %>% head %>% mykable()
```
```{r, label='phos-phos-cdesc'}
phos@cdesc %>% mykable()
```

# Data overview

## filter out not STY
```{r, label='phos-filter-sty'}
.tokeep <- phos@rdesc %>% filter(AA %in% c("S", "T", "Y")) %>% rownames
#.phos %>% cmapR::subset_gct(rid = .tokeep)
.phos <- cmapR::subset_gct(phos, rid = .tokeep, cid=phos@cid)
print(phos@cdesc)
print(phos@cid)
print(head(mat(phos)))
phos <-.phos

```


## distributions { .tabset }
### prof
```{r, label='phos-dist1', eval=F}
prof %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  ggviolin(x="id.y", y="logvalue", add = "median_mad")  %>%
  ggpar(x.text.angle=90)

```

### prof log mednorm
```{r, label='phos-dist2', cache=T, eval=F}
prof %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  group_by(id.y) %>%
  mutate(logvalue_mednorm = logvalue-median(logvalue, na.rm=T)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm", add="mean_sd") %>%
  ggpar(x.text.angle=90)
```

### prof mednorm log
```{r, label='phos-dist3', cache=T, eval=F}
prof %>% melt_gct() %>%
  mutate(value_mednorm = value/median(value, na.rm=T)) %>%
  mutate(logvalue_mednorm_div = log(value_mednorm)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm_div", add="mean_sd") %>%
  ggpar(x.text.angle=90)

```


## distributions - phos { .tabset }




### phos
```{r, label='phos-dist4', cache=TRUE, eval=F}
phos %>% melt_gct() %>% ggviolin(x="id.y", y="value")  %>% ggpar(x.text.angle=90)
```

### phos log
```{r, label='phos-dist5', cache=T, eval=F}
phos %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  ggviolin(x="id.y", y="logvalue", add="median_mad") %>%
  ggpar(x.text.angle=90)
```

### phos log mednorm

```{r, label='phos-dist6', eval=F}
phos %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  group_by(id.y) %>%
  mutate(logvalue_mednorm = logvalue-median(logvalue, na.rm=T)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm", add="mean_sd") %>%
  ggpar(x.text.angle=90)

```

### phos mednorm log
```{r, label='phos-dist7'}
phos %>% melt_gct() %>%
  group_by(id.y) %>%
  mutate(value_mednorm = value/median(value, na.rm=T)) %>%
  mutate(logvalue_mednorm_div = log(value_mednorm)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm_div", add="mean_sd") %>%
  ggpar(x.text.angle=90)

```






## normalize

```{r, label='phos-normalize-expr-matrix-1', echo=T, eval=T}
# prof %>% melt_gct() %>% group_by(id.y) %>% filter(value>0) %>% summarise(median(value))
# prof %>% melt_gct() %>% group_by(id.y) %>% summarise(median(value))

.min_nonzero <- mat(prof)[ mat(prof) > 0 ] %>% min
.min_nonzero<-0
  #mutate(value = (value/.min_nonzero)+.min_nonzero) %>%

.medvals <- prof %>% melt_gct() %>% group_by(id.y) %>%
  summarise(medval = median(value, na.rm=T))

.medvals <- prof %>% melt_gct() %>% filter(value > 0) %>% group_by(id.y) %>%
  summarise(medval = median(value, na.rm=T))

# .quantilevals <- prof %>% melt_gct() %>% filter(value > 0) %>% group_by(id.y) %>%
#   summarise(medval = quantile(value, na.rm=T, probs = .5, names = F))

  #mutate(value_mednorm = (value+.min_nonzero*.01)/median(value, na.rm=T)) %>%
.normalized_prof_matrix <- prof %>%
  melt_gct() %>%
  left_join(.medvals, by = 'id.y') %>%
  group_by(id.y) %>%
    mutate(value_mednorm = value / medval) %>%
  ungroup() %>%
  mutate(value_mednorm_log=log10(value_mednorm)) %>%
  group_by(id.x) %>%
  mutate(zscore = myzscore(value_mednorm_log)) %>%
  ungroup()
  #mutate(value_mednorm_log = value_log-median(value_log, na.rm=T) ) %>%
  #mutate(zscore=scale(value_mednorm_log)) %>%
  #mutate(value_mednorm=value/median(value, na.rm=T)) %>%
 # ggviolin(x='id.y', y='zscore')
# .normalized_prof_matrix

.prof_correct_mat_wide <- prof_correct %>% melt_gct() %>%
  group_by(id.x) %>%
  mutate(zscore=myzscore(value)) %>%
  ungroup() %>%
  pivot_wider(id_cols=id.x,
   values_from=zscore,
   names_from = id.y
   ) %>% as.data.frame()
rownames(.prof_correct_mat_wide) <- .prof_correct_mat_wide$id.x
.prof_correct_mat_wide$id.x <- NULL # delete
.prof_correct_mat_wide <- as.matrix(.prof_correct_mat_wide)

prof_correct_zscore <- new("GCT",
  mat = .prof_correct_mat_wide[prof_correct@rid, prof_correct@cid],
  rid = prof_correct@rid,
  cid = prof_correct@cid,
  rdesc = prof_correct@rdesc,
  cdesc = prof_correct@cdesc
  )


.normalized_prof_matrix_wide_log <- .normalized_prof_matrix %>%
  pivot_wider(id_cols=id.x,
   values_from=value_mednorm_log,
   names_from = id.y
   ) %>% as.data.frame()
rownames(.normalized_prof_matrix_wide_log) <- .normalized_prof_matrix_wide_log$id.x
.normalized_prof_matrix_wide_log$id.x <- NULL # delete
.normalized_prof_matrix_wide_log <- as.matrix(.normalized_prof_matrix_wide_log)


.normalized_prof_matrix_wide_zscore <- .normalized_prof_matrix %>%
  pivot_wider(id_cols=id.x,
   values_from=zscore,
   names_from = id.y
   ) %>% as.data.frame()

rownames(.normalized_prof_matrix_wide_zscore) <- .normalized_prof_matrix_wide_zscore$id.x
.normalized_prof_matrix_wide_zscore$id.x <- NULL # delete
.normalized_prof_matrix_wide_zscore <- as.matrix(.normalized_prof_matrix_wide_zscore)

prof_zscore <- new("GCT",
  mat = .normalized_prof_matrix_wide_zscore[prof@rid, prof@cid],
  rid = prof@rid,
  cid = prof@cid,
  rdesc = prof@rdesc,
  cdesc = prof@cdesc
  )

prof_correct_zscore %>% melt_gct() %>% filter(id.x=="10178909")
prof_zscore %>% melt_gct() %>% filter(id.x=="10178909")
prof_zscore %>% melt_gct() %>% filter(id.x=="10178909") %>% pivot_wider(values_from = value, names_from = id.y, id_cols = id.x)
prof_correct_zscore %>% melt_gct() %>% filter(id.x=="10178909") %>% pivot_wider(values_from = value, names_from = id.y, id_cols = id.x)

prof_log <- new("GCT",
  mat = as.matrix(.normalized_prof_matrix_wide_log)[prof@rid, prof@cid],
  rid = prof@rid,
  cid = prof@cid,
  rdesc = prof@rdesc,
  cdesc = prof@cdesc
  )

```

## heatmap def
```{r, label='phos-heatmap-def'}
make_heatmap <- function(.gct, show_row_names=F, pos_sel=NULL){
  .negsel <- c("sample", "plex", "recno", "id", "runno", "searchno", "assay", "replicate", "name",
               "lu",
               "ProtocolNo",
               "Washes",
               "Date",
               "Experimenter",
               "X1",
               "X1Detail",
               "X2",
               "X2Detail",
               "X1.1",
               "X1Detail.1",
               "X2.1",
               "X2Detail.1",
               "X1.2",
               "X1Detail.2",
               "X2.2",
               "X2Detail.2"
               )
  .column_df <- .gct@cdesc %>% select(-any_of(.negsel))
  if (!is.null(pos_sel))
    .column_df <- .gct@cdesc %>% select(any_of(pos_sel))

  library(ComplexHeatmap)

  recycle_colors <- function(type = "qual", n = 8) {
    # original_palette <- brewer_pal(palette = palette_name)(min(n, 8))
    original_palette <- brewer_pal(type = "qual", palette = 2)(min(n, 8))
    recycled_palette <- rep_len(original_palette, length.out = n)
    return(recycled_palette)
  }

  .color_list <- .column_df %>%
    as.list() %>%
    map(~ recycle_colors(n = length(unique(.))))

  # fix this maybe if better
  for (col in colnames(.column_df)) {
    # print(col)
    # names(.color_list[[col]]) <-  unique(.column_df[[col]])
    names(.color_list[[col]]) <- unique(.column_df[[col]])
  }


.top_annotation <- columnAnnotation(
  df = .column_df %>% as.data.frame(),
  col = .color_list
  # col = anno_block(gp = gpar(fill = .color_list$metavar1))
)

  #mat_zscore <- apply(mat(.gct) %>% head(50), MARGIN=1, myzscore)
  # mat_zscore <- apply(mat(.gct), MARGIN=1, myzscore)
  # dim(mat_zscore)
  # mat_zscore <- t(mat_zscore)  # transpose back to original structure
  # colnames(mat_zscore) <- colnames(mat(.gct))  # Save column names as row names before transpose
  # mat_zscore <- mat_zscore[, rownames(.column_df)]
  #mat_zscore = mat_zscore[, rownames(.column_df)]

  #mat_zscore <- mat_zscore[, rownames(.column_df)] # make sure order is the same
  mat <- mat(.gct)
  mat <- mat[, .gct@cdesc$id] # make sure same order

  .h <- ComplexHeatmap::Heatmap(
    mat,
    top_annotation = .top_annotation,
    clustering_distance_rows = dist_no_na,
    clustering_method_rows = 'ward.D2',
    clustering_method_columns = 'ward.D2',
    clustering_distance_columns = dist_no_na,
    column_title_side = "top",
    column_title_rot = 0,
    column_names_gp = gpar(fontsize = 9),
    row_names_gp = gpar(fontsize=6),
    border = TRUE,
    use_raster = T,
    show_row_names = show_row_names,
    show_row_dend = T,

  )
  .h
}
```


```{r}
cols_to_annotate <- c("CellTissue", "Genotype")
cols_to_annotate <- c("Genotype", "recno")
# make_heatmap(phos_gct_batch, pos_sel = cols_to_annotate)
make_heatmap(phos_gct_zscore, pos_sel = cols_to_annotate)
```



```{r}
prof_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```

```{r}
prof_correct_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```

```{r}
mat(prof_correct_zscore) %>%
  Heatmap(show_row_dend = F, show_row_names = F)
```

```{r}
make_heatmap(prof_correct_zscore)
```


```{r}
make_heatmap(prof_zscore)
```


```{r, label='phos-normalize-expr-matrix2', echo=F, eval=T}
.normalized_phos_matrix <- phos %>% melt_gct() %>%
  group_by(id.y) %>%
  mutate(value_mednorm = value / median(value, na.rm=T)) %>%
  ungroup() %>%
  mutate(logvalue_mednorm = log(value_mednorm)) %>%
  group_by(id.x) %>%
  mutate(zscore=myzscore(logvalue_mednorm))

.normalized_phos_matrix_wide_zscore <- .normalized_phos_matrix %>%
  pivot_wider(id_cols=id.x, values_from=zscore, names_from = id.y ) %>%
  as.data.frame()
rownames(.normalized_phos_matrix_wide_zscore) <- .normalized_phos_matrix_wide_zscore$id.x
.normalized_phos_matrix_wide_zscore$id.x <- NULL # delete

#mat(phos) <- .normalized_phos_matrix %>% as.matrix
.normalized_phos_matrix_wide_log <- .normalized_phos_matrix %>%
  pivot_wider(id_cols=id.x, values_from=logvalue_mednorm, names_from = id.y ) %>%
  as.data.frame()
rownames(.normalized_phos_matrix_wide_log) <- .normalized_phos_matrix_wide_log$id.x
.normalized_phos_matrix_wide_log$id.x <- NULL # delete

phos_zscore <- new("GCT",
  mat = as.matrix(.normalized_phos_matrix_wide_zscore)[phos@rid, phos@cid],
  rid = phos@rid,
  cid = phos@cid,
  rdesc = phos@rdesc,
  cdesc = phos@cdesc
  )

phos_log <- new("GCT",
  mat = as.matrix(.normalized_phos_matrix_wide_log)[phos@rid, phos@cid],
  rid = phos@rid,
  cid = phos@cid,
  rdesc = phos@rdesc,
  cdesc = phos@cdesc
  )

#phos %>% melt_gct() %>% group_by(id.y) %>% mutate(value_mednorm=log(value/median(value,na.rm=T))) %>%
#  ggviolin(x='id.y', y='value_mednorm')
phos_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```




```{r}
make_heatmap(phos_zscore)
```

# heatmaps


## heatmap prof
```{r, label='phos-heatmap-prof', cache=FALSE}
make_heatmap(prof)
```

## heatmap phos
```{r, label='phos-heatmap-phos', cache=FALSE}
#make_heatmap(phos)
print(phos@cdesc)
print(phos@cid)
print(head(mat(phos)))
```

#

# PCA
render PCA.Rmd
```{r, label=phos-pca-render-func, message=T, warning=T, results='show', eval=T, class.source='fold-hide', eval=F}

#.params <- params$PCA

# .params$method <- "kmeans"

render_pca_report <- function(center = T, scale = F, title="PCA", ...) {
  .output_file_name <- paste0("PCA", "title_", title, "center_", center, "scale_", scale, ..., ".html")
  rmarkdown::render("PCA.Rmd",
    output_file = .output_file_name,
    quiet = T,
    params = list(
      center = center,
      scale = scale,
      title = title
    )
    # params=.params,
    # envir = globalenv(),
    # envir = new_env
    # documentation=1,
  )
  return(.output_file_name)
  # c(params$ROOTPATH, .output_file_name) %>% fs::path_join() %>% fs::path_abs()
}
```

```{r, label="phos-pca-make-prof", echo=T, eval=F}
# TODO: allow customization
gct <- prof

.center <- T
.scale <- T

gct <- prof
.out <- render_pca_report(center = .center, scale = T,
                          color = "genotype",
                          title="prof")
```
LINK
```{r, label="pca-report-1-html-pagetest", echo=F, results="asis", eval=F}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

```{r, label="phos-pca-make-phos", echo=T, eval=F}
gct <- phos
.out <- render_pca_report(center = .center, scale = T,
                          color = "genotype",
                          title="phos")

# html
# .html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)

# htmltools::includeHTML(.out)
```

LINK
```{r, label="pca-report-2-html-pagetest", echo=F, results="asis", eval=F}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```



## examine particular sites



```{r, label='phos-makeptable'}
make_p_table <- function(geneid){
  geneid <- geneid
  prof_expression_data <- mat(prof)[geneid,]
  .psites <- phos@rdesc[phos@rdesc$GeneID == geneid,]
  phos_expression_data <- mat(phos)[rownames(.psites), ]
  #phos_site <- phos@rdesc$phosphosites  # Adjust this line based on how phosphosites is stored in your data
  #prof$Phosphosites <- phos_site
  DT::datatable(.psites)
}
```

```{r, label='phos-geneid-selection'}
  geneid <- "38418"
  make_p_table(geneid)

```




# Differential Analysis
## Differential Analysis
###limma

```{r}


mod <- model.matrix(~0 + genotype, phos@cdesc)
colnames(mod) <- make.names(colnames(mod)) #where is this used
limma_input <- mat(phos_log)
limma_input <- limma_input[apply(limma_input, 1, function(row) all(row != -Inf)), ]
limma_input <- limma_input[ ,rownames(mod) ]

# Select rows where all values are not -Inf


contrast_list <- c(
  "genotypeIntS6 - genotypeWT",
  "genotypeIntS12 - genotypeWT",
  "genotypeIntS6 - genotypeIntS12",
  #"genotypeIntS6 + genotypeIntS12"
  )
contrasts <- makeContrasts(contrasts=contrast_list, levels = mod) #need to make variable contrast def

prefit <- lmFit(limma_input, mod)
fit <- contrasts.fit(prefit, contrasts) %>% eBayes(trend = T, robust = T)
```


```{r}
limma_res <- purrr::map_dfr(
  colnames(contrasts),
  ~ topTable(fit, n=Inf, sort.by="none", coef=.x, confint = F  ) %>%
  rownames_to_column("GeneIDSite") %>%
  mutate(contrast=.x)
)
.file <- file.path(ROOTPATH, "MSPC000967_limma_res.tsv")
limma_res %>% write_tsv(.file)
# %>%
  #rownames_to_column("GeneIDSite") %>%
  #tibble() %>%
  #arrange(contrast, P.Value) #move from here to plot topN

#limma_res %>% head(20)
```


## top50 { .tabset }

### ints6 vs wt

```{r, fig.width=10, fig.height=9}
.c <- "genotypeIntS6 - genotypeWT"
.top50 <- limma_res %>%
  filter(contrast == .c) %>%
  arrange(P.Value) %>%
  head(75)


subset_gct(phos_zscore, rid=.top50$GeneIDSite) %>% make_heatmap(show_row_names=T)

```

### ints12 vs wt

```{r, fig.width=10, fig.height=9}
.c <- "genotypeIntS12 - genotypeWT"
.top50 <- limma_res %>%
  filter(contrast == .c) %>%
  arrange(P.Value) %>%
  head(75)


subset_gct(phos_zscore, rid=.top50$GeneIDSite) %>% make_heatmap(show_row_names=T)

```





```{r, fig.width=10, fig.height=9}
.c <- "genotypeIntS6 - genotypeIntS12"
.top50 <- limma_res %>%
  filter(contrast == .c) %>%
  arrange(P.Value) %>%
  head(75)


subset_gct(phos_zscore, rid=.top50$GeneIDSite) %>% make_heatmap(show_row_names=T)

```