---
title: "PHOS"
author: "AlexSaltman"
date: "`r Sys.Date()`"
# runtime: shiny
output:
  html_document:
    code_folding: hide
    highlight: zenburn # pygments, kate, monochrome, zenburn
    toc: yes
params:
  use_paramfile:
    value: TRUE
    description: "Boolean value whether or not to load param file if present"
  paramfile:
    value: "../params/PHOS.yaml"
  ROOTPATH:
    value: "/media/alex/20230612/MSPC000936/MSPC000936/"
  PROF:
    #value: "MSPC000967_TMT10_prof_Config_data_gct_linear_all_1more_nonzero_normtype_none_n10x7507.gct"
    value: "MSPC000967_TMT10_prof_Config_data_gct_all_1more_nonzero_normtype_median_n10x7507.gct"
    description: "gct file for one profiling experiment"
  PROF_PATH:
    value: "./exports/prof"
    description: "directory location for profiling gct files"
  PHOS:
    #value: "51963_1_6_site_table_nr.tsv_n10x14785.gct"
    #description: "gct file for one phosphoprofiling experiment"
  PHOS_PATH:
    value: "IMAC/"
    description: "directory location for phos gct files"
  cluster_methods:
    value: [kmeans]
  formula:
    value: "~0+Extract_Genotype:Extract_Treatment"
    description: "will move to separate volcano.Rmd document in the future, probably"
  pca:
    value:
      color:
        value: Extract_Genotype
      marker:
        value: Extract_Treatment
      annotate:
        value: true
      center:
        value: true
      scale:
        value:
          true
  cluster:
    value:
      cluster_rows:
        value: true
        description: "cluster the rows"
      cluster_columns:
        value: true
      cluster_row_slices:
        value: true
      cluster_col_slices:
        value: true
      clustering_distance_rows:
        value: euclidean
      clustering_method_rows:
        value: ward.d2
      legend_include:
        value:
        description: "metadata variables to include"
      legend_exclude:
        value:
        description: "metadata variables to exclude"
  logging:
    value:
      message: TRUE
      warning: TRUE
      quiet: TRUE
end:
#

---


```{r label=phos.setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(cmapR))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(fs))
suppressPackageStartupMessages(library(limma))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
suppressPackageStartupMessages(library(tidymodels))
library(tidymodels)
library(tibble)
library(magrittr)
source("funcs.R")

mykable <- function(data) {
    knitr::kable(data, booktabs = TRUE, digits = 4) %>%
        kable_styling(latex_options = c("striped", "scale_down"),
                      bootstrap_options = c("striped", 'condensed', "bordered")
                      )
}

get_param <- function(param){
  if(is.list(param) && "value" %in% names(param)){
    return(param$value)
  } else{
    return(param)
  }
}
```

# Load

## params

```{r, label=phos.load.params, class.source='fold-show'}

#.f <- params$paramfile
.f <- get_param(params$paramfile)
.use_paramfile <- get_param(params$use_paramfile)
if (!is.null(.f) && fs::file_exists(.f) && params$use_paramfile == TRUE) {
  subparams <- yaml::yaml.load_file(.f)
  cat(paste("loading", .f))
  #params <- modifyList(params, local_params)
} else{
  subparams <- params
}
# cat(paste(subparams))
# (params)
# (subparams)
```

```{r, label=phos.set.rootdir, class.source="fold-show", eval=F} 
# # not even sure we need this
# ROOTPATH <- get_param(subparams$ROOTPATH)
# ROOTPATH <- fs::path_abs(ROOTPATH) 
# (ROOTPATH)
# setwd(ROOTPATH)
```

## gct

```{r, label="phos.load.data", class.source='fold-show'}
# ROOTPATH <- fs::path_abs(subparams$ROOTPATH)
# setwd(ROOTPATH)

ROOTPATH <- get_param(subparams$ROOTPATH)
PROF_PATH <- fs::path(ROOTPATH, get_param(subparams$PROF_PATH))
PHOS_PATH <- fs::path(ROOTPATH, get_param(subparams$PHOS_PATH))
# PHOS_PATH <- fs::path_abs(subparams$PHOS_PATH)

#prof_files <- PROF_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")

prof_files <- PROF_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")
(prof_files)

phos_files <- PHOS_PATH %>% fs::dir_ls(recurse=T, regexp = "*.gct")
(phos_files)

parse_gct_dropzeros <- function (file){
  gct <- cmapR::parse_gctx(file)
  mat <- gct@mat
  mat_nonzero <- mat[apply(mat, 1, sum) != 0, ]
  remaining_rids <- rownames(mat_nonzero)
  gct <- gct %>% subset_gct(rid=remaining_rids)
  # this has to go somewhere else and be more costomizable
  # valid_cids <- gct@cdesc %>% filter(Extract_Genotype != "RefMix") %>% pull(id)
  # gct <- gct %>% subset_gct(cid = valid_cids)
}
prof_gcts <- prof_files %>% purrr::map(parse_gct_dropzeros)

#prof_gcts <- prof_files %>% purrr::map(cmapR::parse_gctx)
phos_gcts <- phos_files %>% purrr::map(cmapR::parse_gctx)
# need to remove sample identifying info that is in the rdesc
cols_to_drop <- c("recno", "runno", "searchno", "basename", "TopScore")
phos_gcts <- purrr::map(phos_gcts, function(gct) {
  gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
  return(gct)
})

# need to remove sample identifying info that is in the rdesc
cols_to_drop <- c("recno", "runno", "searchno", "basename", "GPGroup", "PSMs", "PeptidePrint", "EXPRecNo", "EXPRunNo", "EXPSearchNo", "LabelFLAG")
prof_gcts <- purrr::map(prof_gcts, function(gct) {
  gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
  return(gct)
})



#gct@rdesc <- gct@rdesc[, !colnames(gct@rdesc) %in% cols_to_drop]
```


## filter 
```{r, label=phos.filter}
#phos_gcts %<>% purrr::map(~cmapR::subset_gct())
# filter out not STY aa modi


phos_gcts_f <- phos_gcts %>%
  purrr::map(~cmapR::subset_gct(., rid = .@rdesc%>%dplyr::filter(AA %in% c("S", "T", "Y")) %>% rownames)  )

phos_gct_merged <- base::Reduce(accumulate=F,
                    f=function(...) cmapR::merge_gct(..., dim="col"),
                    x=phos_gcts_f
)
rownames(phos_gct_merged@cdesc) <- phos_gct_merged@cdesc$id # always called ID when melting, taken from rownames

cidcounts <- phos_gct_merged %>% melt_gct %>% group_by(recno, runno, searchno) %>% summarise(count=n_distinct(id.y)) %>% ungroup
ridcounts <- phos_gct_merged %>% melt_gct %>% group_by(id.x) %>% summarise(count=n_distinct(id.y)) %>% ungroup
.maxval <- 30
.maxval <- max(ridcounts$count)
rids_filtered <- ridcounts %>% filter(count >= .maxval) %>% pull(id.x)
# here filter for species

phos_gct <- subset_gct(phos_gct_merged, rid=rids_filtered)


prof_gct_merged <- base::Reduce(
  f = function(...) cmapR::merge_gct(..., dim = "col"),
  x=prof_gcts
)
ridcounts <- prof_gct_merged %>% melt_gct %>% group_by(id.x) %>% summarise(count=n_distinct(id.y)) %>% ungroup
.maxval <- max(ridcounts$count)
rids_filtered <- ridcounts %>% filter(count >= .maxval) %>% pull(id.x)
.species_filter <- prof_gct_merged@rdesc %>% filter(TaxonID=="9606")
rids_filtered <- intersect(rids_filtered, .species_filter$GeneID)


# here filter for species
prof_gct <- subset_gct(prof_gct_merged, rid=rids_filtered)
```

## prettify metadata for better display
maybe move this up, but maybe not
```{r, label=phos.prettify.meta, include=T}
# one (of hopefully no more) instance where we modify in place
.display_name <- phos_gct@cdesc$DisplayID
rownames(phos_gct@cdesc) <- .display_name
colnames(phos_gct@mat) <- .display_name
phos_gct@cid <- .display_name
phos_gct@cdesc$id <- .display_name
.new_rdesc <- phos_gct@rdesc %>%
  mutate(GeneSymbol = ifelse(GeneSymbol == "", GeneID, GeneSymbol)) %>% 
  mutate(DisplayName = stringr::str_c(GeneSymbol, "_", ModiType, AApos_Selected, "_", Site, sep = ""))

# this is also where renaming the geneids may occur
# similarly
# .new_gene_names <- make_new_names(rownames(phos_gct@mat))
# rownames(phos_gct@mat) <- .display_name
# phos_gct@rid <- .display_name
.new_gene_names <- .new_rdesc$DisplayName
.count <- .new_rdesc %>% group_by(DisplayName) %>% mutate(n=n()) %>% arrange(-n)
.maxcount <-  max(.count$n) 
if (.maxcount > 1){
  # this is bad
  1+1
}
phos_gct@rdesc <- .new_rdesc

rownames(phos_gct@mat) <- .new_gene_names
phos_gct@rid <- .new_gene_names
phos_gct@rdesc$id <- .new_gene_names # must change this as well or we will get a name mismatch


.display_name <- prof_gct@cdesc$DisplayID
rownames(prof_gct@cdesc) <- .display_name
colnames(prof_gct@mat) <- .display_name
prof_gct@cid <- .display_name
prof_gct@cdesc$id <- .display_name

# .new_gene_names$name
#.mat <- phos_gct@mat
#phos_gct@mat <- .mat
#.cdesc <- phos_gct@cdesc
#.rdesc <- phos_gct@rdesc
#rownames(.cdesc) <- .display_name
#colnames(.mat) <- .display_name



# .phos_gct <- phos_gct %>% 
#   cmapR::GCT(mat = .mat, rid = .@rid, cid = .@cid, cdesc = .cdesc, rdesc = .rdesc)
# .phos_gct@mat %>% colnames %>% head
# .mat %>% colnames %>% head
# rownames(phos_gct@cdesc) <- .display_name
#colnames(phos_gct@mat) <- .display_name
#phos_gct <- 
#phos_gct_batch <- phos_gct %>% cmapR::GCT(mat = mat_combat, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
```

## normalize 
```{r, label="phos.normalize", eval=T}
dfm <- phos_gct %>% melt_gct()
dfm %<>% 
  mutate(value = value / min(value[value != 0], na.rm=T)) %>% 
  group_by(id.y) %>%
  mutate(mednorm = value / median(value, na.rm=TRUE) + 1) %>%
  mutate(logmednorm=log(mednorm)) %>%
  ungroup()

mat_log <- dfm %>% pivot_wider(id_cols=id.x, names_from=id.y, values_from=logmednorm ) %>% as.data.frame
rownames(mat_log) <- mat_log$id.x
mat_log$id.x <- NULL

phos_gct_log <- phos_gct %>% cmapR::GCT(mat = as.matrix(mat_log), rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
#colnames(mat)


dfm <- prof_gct %>% melt_gct()
dfm %<>% 
  mutate(value = value / min(value[value != 0], na.rm=T)) %>% 
  group_by(id.y) %>%
  mutate(mednorm = value / median(value, na.rm=TRUE) + 1) %>%
  mutate(logmednorm=log(mednorm)) %>%
  ungroup()

mat_log <- dfm %>% pivot_wider(id_cols=id.x, names_from=id.y, values_from=logmednorm ) %>% as.data.frame
rownames(mat_log) <- mat_log$id.x
mat_log$id.x <- NULL
prof_gct_log <- prof_gct %>% cmapR::GCT(mat = as.matrix(mat_log), rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)

# add logic for checking if we are to do batch correction
# if batch:
```
## batch correction
```{r, label=phos.batch}

do_batch <- get_param(subparams$batchcorrection$do_batch_correction)
batchvar <- get_param(subparams$batchcorrection$batchvar)
#if 
#this is not needed
batchvalues <- phos_gct_log@cdesc[[batchvar]] %>% as.character()
mat_combat <- sva::ComBat(phos_gct_log@mat, batchvalues)
phos_gct_batch <- phos_gct %>% cmapR::GCT(mat = mat_combat, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
phos_gct_log <- phos_gct_batch


#batchinfo <- prof_gct@cdesc %>% mutate(id=factor(id, levels=colnames(mat_log), ordered=TRUE)) %>% arrange(id)
#batchinfo <- prof_gct_log@cdesc[[batchvar]] %>% as.character()
batchvalues <- prof_gct_log@cdesc[[batchvar]] %>% as.character()
mat_combat <- sva::ComBat(prof_gct_log@mat, batchvalues)
prof_gct_batch <- prof_gct_log %>% cmapR::GCT(mat = mat_combat, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
prof_gct_log <- prof_gct_batch

#else:
#phos_gct_log <- phos_gct %>% cmapR::GCT(mat = mat_log, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
# if not true

```

```{r, label=phos.zscore}

.tmp <- phos_gct_log %>%
  melt_gct() %>%
  group_by(id.x) %>%
  mutate(zscore = myzscore(value)) %>%
  ungroup()
mat_zscore <- .tmp %>%  pivot_wider(id_cols=id.x, names_from=id.y, values_from=zscore) %>% as.data.frame
rownames(mat_zscore) <- mat_zscore$id.x
mat_zscore$id.x <- NULL
mat_zscore <- as.matrix(mat_zscore)

phos_gct_zscore <- phos_gct_log %>% cmapR::GCT(mat = mat_zscore, rid = .@rid, cid = .@cid, cdesc = .@cdesc, rdesc = .@rdesc)
```



```{r, label='phos.manual-metadata', eval=F}
.res<-left_join(prof@rdesc,
          distinct(phos@rdesc, GeneID, GeneSymbol),
          by = c('rdesc' = 'GeneID')
          )
prof@rdesc <- .res

.phos_metadata <- tibble::tribble(
~id, ~name,	~recno,	~runno,	~searchno,	~label,	~assay,	~genotype,	~replicate,
"51963_1_6_126", "DL1-1",	  51963,	1,	6,	"126", 	"IMAC",	"WT",	1,
"51963_1_6_127_N", "DL1-2",	  51963,	1,	6,	"127N",	"IMAC",	"WT",	2,
"51963_1_6_127_C", "DL1-3",	  51963,	1,	6,	"127C",	"IMAC",	"WT",	3,
"51963_1_6_128_N", "IntS6-1",	51963,	1,	6,	"128N",	"IMAC",	"IntS6",	1,
"51963_1_6_128_C", "IntS6-2",	51963,	1,	6,	"128C",	"IMAC",	"IntS6",	2,
"51963_1_6_129_N", "IntS6-3",	51963,	1,	6,	"129N",	"IMAC",	"IntS6",	3,
"51963_1_6_129_C", "IntS6-4",	51963,	1,	6,	"129C",	"IMAC", "IntS6", 	4,
"51963_1_6_130_N", "IntS12-1",	51963, 	1, 	6,	"130N",	"IMAC", "IntS12",	1,
"51963_1_6_130_C", "IntS12-2",	51963, 	1, 	6,	"130C",	"IMAC", "IntS12",	2,
"51963_1_6_131_N", "IntS12-3",	51963, 	1, 	6,	"131",	"IMAC", "IntS12",	3
)
#.phos_metadata$id <- .phos_metadata$name
.phos_sample_metadata <- .phos_metadata %>% mutate(id = factor(id, levels=phos@cid, ordered=T)) %>% arrange(id) %>% as.data.frame()
rownames(.phos_sample_metadata) <- .phos_sample_metadata$id
#phos@cdesc <- .phos_sample_metadata
#rownames(.phos_metadata) <- .phos_metadata$name #
#.cols <- c("id", "recno", "runno", "searchno", "label", "")
#out <- .phos_metadata %>% left_join(prof@cdesc, by = "id")
.the_order_we_want <- colnames(mat(phos))
.newnames <- .phos_sample_metadata %>%
   mutate(id=factor(id, levels(.the_order_we_want), ordered = T)) %>%
   arrange(id) %>%
   pull(name)
rownames(.phos_sample_metadata) <- .newnames
.phos_sample_metadata$id <- rownames(.phos_sample_metadata)
# source('funcs.R')
#colnames(mat(phos)) <- .newnames
#phos@cid <- .newnames
.new <- mat(phos)
colnames(.new) <- .newnames
# .new %>% dim
# phos %>% mat %>% dim
# phos@rdesc %>% dim
.rdesc <- phos@rdesc
.rid <- phos@rid

.phos <- new("GCT",
    mat=.new,
    cdesc=.phos_sample_metadata,
    cid=rownames(.phos_sample_metadata),
    #rid=phos@rid,
    #rid=rownames(.new),
    rid=.rid,
    rdesc=.rdesc,
    )
# subset_gct(phos)
# subset_gct(.phos)
phos <- .phos

#phos <- make_new_gct(phos, mat=.new, cdesc=.phos_metadata, cid=.phos_metadata$name)
#phos@cdesc <- .phos_metadata
# phos2 <- make_new_gct(phos, .new)
# phos <- phos2
```

```{r, label="phos-phos-debug", eval=F}

print(phos@cdesc)
print(phos@cid)
print(head(mat(phos)))
```


## preview metadata
 prof

```{r, label="phos-prof-mat", eval=F}
mat(prof) %>% head %>% mykable()
```

```{r, label="phos-prof-cdesc", eval=F}
prof@cdesc %>% mykable()
```



```{r, label="phos-phos-mat", eval=T}
mat(phos_gct) %>% head %>% mykable()
```
```{r, label='phos-phos-cdesc'}
phos_gct@cdesc %>% mykable()
```

# Data overview

## distributions { .tabset }
### prof
```{r, label='phos-dist1', eval=F}
prof %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  ggviolin(x="id.y", y="logvalue", add = "median_mad")  %>%
  ggpar(x.text.angle=90)

```

### prof log mednorm
```{r, label='phos-dist2', cache=T, eval=F}
prof %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  group_by(id.y) %>%
  mutate(logvalue_mednorm = logvalue-median(logvalue, na.rm=T)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm", add="mean_sd") %>%
  ggpar(x.text.angle=90)
```

### prof mednorm log
```{r, label='phos-dist3', cache=T, eval=F}
prof %>% melt_gct() %>%
  mutate(value_mednorm = value/median(value, na.rm=T)) %>%
  mutate(logvalue_mednorm_div = log(value_mednorm)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm_div", add="mean_sd") %>%
  ggpar(x.text.angle=90)

```


## distributions - phos { .tabset }




### phos
```{r, label='phos-dist4', cache=TRUE, eval=T}
phos_gct_log %>% melt_gct() %>% ggviolin(x="id.y", y="value")  %>% ggpar(x.text.angle=90)
```

### phos log
```{r, label='phos-dist5', cache=T, eval=F}
phos %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  ggviolin(x="id.y", y="logvalue", add="median_mad") %>%
  ggpar(x.text.angle=90)
```

### phos log mednorm

```{r, label='phos-dist6', eval=F}
phos %>% melt_gct() %>%
  mutate(logvalue = log(value)) %>%
  group_by(id.y) %>%
  mutate(logvalue_mednorm = logvalue-median(logvalue, na.rm=T)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm", add="mean_sd") %>%
  ggpar(x.text.angle=90)

```

### phos mednorm log
```{r, label='phos-dist7', eval=F}
phos %>% melt_gct() %>%
  group_by(id.y) %>%
  mutate(value_mednorm = value/median(value, na.rm=T)) %>%
  mutate(logvalue_mednorm_div = log(value_mednorm)) %>%
  ggviolin(x="id.y", y="logvalue_mednorm_div", add="mean_sd") %>%
  ggpar(x.text.angle=90)

```







## normalize

```{r, label='phos-normalize-expr-matrix-1', echo=T, eval=F}
# prof %>% melt_gct() %>% group_by(id.y) %>% filter(value>0) %>% summarise(median(value))
# prof %>% melt_gct() %>% group_by(id.y) %>% summarise(median(value))

.min_nonzero <- mat(prof)[ mat(prof) > 0 ] %>% min
.min_nonzero<-0
  #mutate(value = (value/.min_nonzero)+.min_nonzero) %>%

.medvals <- prof %>% melt_gct() %>% group_by(id.y) %>%
  summarise(medval = median(value, na.rm=T))

.medvals <- prof %>% melt_gct() %>% filter(value > 0) %>% group_by(id.y) %>%
  summarise(medval = median(value, na.rm=T))

# .quantilevals <- prof %>% melt_gct() %>% filter(value > 0) %>% group_by(id.y) %>%
#   summarise(medval = quantile(value, na.rm=T, probs = .5, names = F))

  #mutate(value_mednorm = (value+.min_nonzero*.01)/median(value, na.rm=T)) %>%
.normalized_prof_matrix <- prof %>%
  melt_gct() %>%
  left_join(.medvals, by = 'id.y') %>%
  group_by(id.y) %>%
    mutate(value_mednorm = value / medval) %>%
  ungroup() %>%
  mutate(value_mednorm_log=log10(value_mednorm)) %>%
  group_by(id.x) %>%
  mutate(zscore = myzscore(value_mednorm_log)) %>%
  ungroup()
  #mutate(value_mednorm_log = value_log-median(value_log, na.rm=T) ) %>%
  #mutate(zscore=scale(value_mednorm_log)) %>%
  #mutate(value_mednorm=value/median(value, na.rm=T)) %>%
 # ggviolin(x='id.y', y='zscore')
# .normalized_prof_matrix

.prof_correct_mat_wide <- prof_correct %>% melt_gct() %>%
  group_by(id.x) %>%
  mutate(zscore=myzscore(value)) %>%
  ungroup() %>%
  pivot_wider(id_cols=id.x,
   values_from=zscore,
   names_from = id.y
   ) %>% as.data.frame()
rownames(.prof_correct_mat_wide) <- .prof_correct_mat_wide$id.x
.prof_correct_mat_wide$id.x <- NULL # delete
.prof_correct_mat_wide <- as.matrix(.prof_correct_mat_wide)

prof_correct_zscore <- new("GCT",
  mat = .prof_correct_mat_wide[prof_correct@rid, prof_correct@cid],
  rid = prof_correct@rid,
  cid = prof_correct@cid,
  rdesc = prof_correct@rdesc,
  cdesc = prof_correct@cdesc
  )


.normalized_prof_matrix_wide_log <- .normalized_prof_matrix %>%
  pivot_wider(id_cols=id.x,
   values_from=value_mednorm_log,
   names_from = id.y
   ) %>% as.data.frame()
rownames(.normalized_prof_matrix_wide_log) <- .normalized_prof_matrix_wide_log$id.x
.normalized_prof_matrix_wide_log$id.x <- NULL # delete
.normalized_prof_matrix_wide_log <- as.matrix(.normalized_prof_matrix_wide_log)


.normalized_prof_matrix_wide_zscore <- .normalized_prof_matrix %>%
  pivot_wider(id_cols=id.x,
   values_from=zscore,
   names_from = id.y
   ) %>% as.data.frame()

rownames(.normalized_prof_matrix_wide_zscore) <- .normalized_prof_matrix_wide_zscore$id.x
.normalized_prof_matrix_wide_zscore$id.x <- NULL # delete
.normalized_prof_matrix_wide_zscore <- as.matrix(.normalized_prof_matrix_wide_zscore)

prof_zscore <- new("GCT",
  mat = .normalized_prof_matrix_wide_zscore[prof@rid, prof@cid],
  rid = prof@rid,
  cid = prof@cid,
  rdesc = prof@rdesc,
  cdesc = prof@cdesc
  )

prof_correct_zscore %>% melt_gct() %>% filter(id.x=="10178909")
prof_zscore %>% melt_gct() %>% filter(id.x=="10178909")
prof_zscore %>% melt_gct() %>% filter(id.x=="10178909") %>% pivot_wider(values_from = value, names_from = id.y, id_cols = id.x)
prof_correct_zscore %>% melt_gct() %>% filter(id.x=="10178909") %>% pivot_wider(values_from = value, names_from = id.y, id_cols = id.x)

prof_log <- new("GCT",
  mat = as.matrix(.normalized_prof_matrix_wide_log)[prof@rid, prof@cid],
  rid = prof@rid,
  cid = prof@cid,
  rdesc = prof@rdesc,
  cdesc = prof@cdesc
  )

```

## heatmap def
```{r, label='phos-heatmap-def', include=T, eval=T}
make_heatmap <- function(.gct, show_row_names=F, cut_by=NULL, pos_sel=NULL){
  .negsel <- c("sample", "plex", "recno",
               "runno", "searchno", "assay", "replicate", "name",
               "lu",
               "ProtocolNo",
               "Washes",
               "Date",
               "Experimenter",
               "Separation_1",
               "Separation_1Detail",
               "Separation_2",
               "Separation_2Detail",
               # "X1.1",
               # "X1Detail.1",
               # "X2.1",
               # "X2Detail.1",
               # "X1.2",
               # "X1Detail.2",
               # "X2.2",
               # "Separation_2Detail"
               "Exp_Date",
               "Exp_Experimenter",
               "Exp_ExpType",
               "EXPRunNo",
               "ExpSearchNo"
               )
  .column_df <- .gct@cdesc %>% select(-any_of(.negsel))
  

  column_split <- NULL
  if (!is.null(cut_by)) {
    for (.col in cut_by) {
      .column_df[[.col]] <- factor(.column_df[[.col]], ordered = TRUE)
    }
    #.column_df <- .column_df %>% arrange(!!cut_by)
    column_split <- lapply(.column_df[cut_by], as.factor)
  } 

  recycle_colors <- function(type = "qual", n = 8) {
    # original_palette <- brewer_pal(palette = palette_name)(min(n, 8))
    original_palette <- brewer_pal(type = "qual", palette = 2)(min(n, 8))
    recycled_palette <- rep_len(original_palette, length.out = n)
    return(recycled_palette)
  }

  # note we do not have the extra columns anymore
  .nocolor <- c("id")
  if (!is.null(pos_sel)) .cols <- pos_sel
  else .cols <- colnames(.column_df)
  .color_list <- .column_df[, .cols] %>%
    select(-any_of(.nocolor)) %>% 
    as.list() %>%
    map(~ recycle_colors(n = length(unique(.))))

  # fix this maybe if better
  for (col in names(.color_list)) {
    # print(col)
    # names(.color_list[[col]]) <-  unique(.column_df[[col]])
    names(.color_list[[col]]) <- unique(.column_df[[col]])
  }


  .top_annotation <- columnAnnotation(
    df = .column_df[ , names(.color_list) ] %>% as.data.frame(),
    col = .color_list
    # col = anno_block(gp = gpar(fill = .color_list$metavar1))
  )

  #mat_zscore <- apply(mat(.gct) %>% head(50), MARGIN=1, myzscore)
  # mat_zscore <- apply(mat(.gct), MARGIN=1, myzscore)
  # dim(mat_zscore)
  # mat_zscore <- t(mat_zscore)  # transpose back to original structure
  # colnames(mat_zscore) <- colnames(mat(.gct))  # Save column names as row names before transpose
  # mat_zscore <- mat_zscore[, rownames(.column_df)]
  #mat_zscore = mat_zscore[, rownames(.column_df)]

  #mat_zscore <- mat_zscore[, rownames(.column_df)] # make sure order is the same
  mat <- mat(.gct)
  mat <- mat[, .gct@cdesc$id] # make sure same order
  mat <- mat[, .column_df$id] # make sure same order

  .h <- ComplexHeatmap::Heatmap(
    mat,
    top_annotation = .top_annotation,
    column_split = column_split,
    clustering_distance_rows = dist_no_na,
    clustering_method_rows = 'ward.D2',
    clustering_method_columns = 'ward.D2',
    clustering_distance_columns = dist_no_na,
    column_title_side = "top",
    column_title_rot = 0,
    column_names_gp = gpar(fontsize = 9),
    row_names_gp = gpar(fontsize=6),
    border = TRUE,
    use_raster = T,
    show_row_names = show_row_names,
    show_row_dend = T,
  )
  .h
}
```


```{r, label="phos.phos_alldata_heatmap", eval=T}
cols_to_annotate <- c("CellTissue", "Genotype")
cols_to_annotate <- c("Genotype", "recno")
# make_heatmap(phos_gct_batch, pos_sel = cols_to_annotate)
make_heatmap(phos_gct_zscore, pos_sel = cols_to_annotate)
```


```{r, label="phos.somemplot", eval=F}
prof_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```

```{r, label="phos.someotherplot", eval=F}
prof_correct_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```

```{r, eval=F}
mat(prof_correct_zscore) %>%
  Heatmap(show_row_dend = F, show_row_names = F)
```

```{r, eval=F}
make_heatmap(prof_correct_zscore)
```


```{r, eval=F}
make_heatmap(prof_zscore)
```


```{r, label='phos-normalize-expr-matrix2', echo=F, eval=F}
.normalized_phos_matrix <- phos %>% melt_gct() %>%
  group_by(id.y) %>%
  mutate(value_mednorm = value / median(value, na.rm=T)) %>%
  ungroup() %>%
  mutate(logvalue_mednorm = log(value_mednorm)) %>%
  group_by(id.x) %>%
  mutate(zscore=myzscore(logvalue_mednorm))

.normalized_phos_matrix_wide_zscore <- .normalized_phos_matrix %>%
  pivot_wider(id_cols=id.x, values_from=zscore, names_from = id.y ) %>%
  as.data.frame()
rownames(.normalized_phos_matrix_wide_zscore) <- .normalized_phos_matrix_wide_zscore$id.x
.normalized_phos_matrix_wide_zscore$id.x <- NULL # delete

#mat(phos) <- .normalized_phos_matrix %>% as.matrix
.normalized_phos_matrix_wide_log <- .normalized_phos_matrix %>%
  pivot_wider(id_cols=id.x, values_from=logvalue_mednorm, names_from = id.y ) %>%
  as.data.frame()
rownames(.normalized_phos_matrix_wide_log) <- .normalized_phos_matrix_wide_log$id.x
.normalized_phos_matrix_wide_log$id.x <- NULL # delete

phos_zscore <- new("GCT",
  mat = as.matrix(.normalized_phos_matrix_wide_zscore)[phos@rid, phos@cid],
  rid = phos@rid,
  cid = phos@cid,
  rdesc = phos@rdesc,
  cdesc = phos@cdesc
  )

phos_log <- new("GCT",
  mat = as.matrix(.normalized_phos_matrix_wide_log)[phos@rid, phos@cid],
  rid = phos@rid,
  cid = phos@cid,
  rdesc = phos@rdesc,
  cdesc = phos@cdesc
  )

#phos %>% melt_gct() %>% group_by(id.y) %>% mutate(value_mednorm=log(value/median(value,na.rm=T))) %>%
#  ggviolin(x='id.y', y='value_mednorm')
phos_zscore %>% melt_gct() %>% ggviolin(x='id.y', y='value')
```




# heatmaps


## heatmap prof
```{r, label='phos-heatmap-prof', cache=FALSE, eval=F}
make_heatmap(prof)
```

## heatmap phos
```{r, label='phos-heatmap-phos', cache=FALSE, eval=F}
#make_heatmap(phos)
print(phos@cdesc)
print(phos@cid)
print(head(mat(phos)))
```



# PCA
## phos
render PCA.Rmd
```{r, label=phos-pca-render-func, message=T, warning=T, results='show', eval=T, class.source='fold-hide', eval=T}

#.params <- params$PCA

# .params$method <- "kmeans"

render_pca_report <- function(center = T, scale = F, title="PCA", color=NULL, marker=NULL, ...) {
  .output_file_name <- paste0("PCA", "title_", title, "center_", center, "scale_", scale, ".html")
  #.output_file_name <- fs::path(params$ROOTPATH, .output_file_name)
  
  # Capture additional arguments into a list
  additional_params <- list(...)

  rmarkdown::render("PCA.Rmd",
    output_file = .output_file_name,
    quiet = T,
    params = c(list(
      center = center,
      scale = scale,
      title = title,
      color = color,
      marker = marker,
      use_paramfile = F
      ),
      additional_params
    )
    # params=.params,
    # envir = globalenv(),
    # envir = new_env
    # documentation=1,
  )
  return(.output_file_name)
  # c(params$ROOTPATH, .output_file_name) %>% fs::path_join() %>% fs::path_abs()
}
```

```{r, label="phos.pca.phos.scaled", echo=T, eval=T}
# TODO: allow customization
gct <- phos_gct_log
#gct <- prof

.center <- T
.scale <- T
.pca_params <- get_param(subparams$pca)
# get subset pca params 
# there is no check here to see if it is defined in the yaml file, 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
.annotate <- .pca_params$annotate %>% get_param()
#.annotate <- T

.out <- render_pca_report(center = .center,
                          scale = .scale,
                          title="PCA_phospho",
                          color = .color,
                          marker = .marker,
                          annotate = .annotate
)
```
LINK
```{r, label="pca.report.1.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

```{r, label="phos.pca.phos.noscaled", echo=T, eval=T}
gct <- phos_gct_log

.center <- T
.scale <- F
.pca_params <- get_param(subparams$pca)
# get subset pca params 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
# .annotate <- .pca_params$annotate %>% get_param() # why is this n ot working?
.annotate <- T


.out <- render_pca_report(center = .center,
                          scale = .scale,
                          color = .color,
                          marker = .marker,
                          annotate = .annotate,
                          title="PCA_phospho"
                          )

# html
# .html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)

# htmltools::includeHTML(.out)
```

LINK
```{r, label="pca.report.2.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

## prof

```{r, label="phos.pca.prof.scaled", echo=T, eval=T}
# TODO: allow customization
gct <- prof_gct_log
#gct <- prof

.center <- T
.scale <- T
.pca_params <- get_param(subparams$pca)
# get subset pca params 
# there is no check here to see if it is defined in the yaml file, 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
.annotate <- .pca_params$annotate %>% get_param()
#.annotate <- T

.out <- render_pca_report(center = .center,
                          scale = .scale,
                          title="PCA_profiling",
                          color = .color,
                          marker = .marker,
                          annotate = .annotate
)
```
LINK
```{r, label="pca.report.3.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

```{r, label="phos.pca.prof.noscaled", echo=T, eval=T}
gct <- prof_gct_log

.center <- T
.scale <- F
.pca_params <- get_param(subparams$pca)
# get subset pca params 
.color <- .pca_params$color %>% get_param()
.marker <- .pca_params$marker %>% get_param()
# .annotate <- .pca_params$annotate %>% get_param() # why is this n ot working?
.annotate <- T


.out <- render_pca_report(center = .center,
                          scale = .scale,
                          title = "PCA_profiling",
                          color = .color,
                          marker = .marker,
                          annotate = .annotate
                          )

# html
# .html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)

# htmltools::includeHTML(.out)
```

LINK
```{r, label="pca.report.4.link", echo=F, results="asis", eval=T}
.html <- cat(paste0("[", .out, "](", .out, ")"), "\n\n")
# print(.html)
## {-}
```

## examine particular sites



```{r, label='phos-makeptable', eval=F}
make_p_table <- function(geneid){
  geneid <- geneid
  prof_expression_data <- mat(prof)[geneid,]
  .psites <- phos@rdesc[phos@rdesc$GeneID == geneid,]
  phos_expression_data <- mat(phos)[rownames(.psites), ]
  #phos_site <- phos@rdesc$phosphosites  # Adjust this line based on how phosphosites is stored in your data
  #prof$Phosphosites <- phos_site
  DT::datatable(.psites)
}
```

```{r, label='phos-geneid-selection', eval=F}
  geneid <- "38418"
  make_p_table(geneid)

```




# Differential Analysis
## Differential Analysis
###limma

```{r eval=T}

formula <- get_param(subparams$formula) %>% as.formula()
( contrasts_str <- get_param(subparams$contrasts) )
( contrasts_list <- strsplit(contrasts_str, "\n")[[1]] ) # strsplit returns vector array length 1

#(  contrasts2 <- sub("^['\"]|['\"]$", "", contrasts) )
( contrasts_list <- purrr::map(contrasts_list, 
                               ~gsub("^['\"]|['\"]$", "", .) )
                               )

#mod <- model.matrix(~0 + genotype, phos@cdesc)
mod <- model.matrix(formula, phos_gct_log@cdesc)

colnames(mod) <- make.names(colnames(mod)) #where is this used
limma_input <- phos_gct_log %>% mat()
# limma_input <- limma_input[apply(limma_input, 1, function(row) all(row != -Inf)), ]
limma_input <- limma_input[ ,rownames(mod) ]

# Select rows where all values are not -Inf


#contrast_list <- c(
#  "genotypeIntS6 - genotypeWT",
#  "genotypeIntS12 - genotypeWT",
#  "genotypeIntS6 - genotypeIntS12",
#  #"genotypeIntS6 + genotypeIntS12"
#  )
# contrasts <- makeContrasts(contrasts=contrast_list, levels = mod) #need to make variable contrast def
contrasts <- makeContrasts(contrasts = contrasts_list, levels = mod)
prefit <- lmFit(limma_input, mod)
fit <- contrasts.fit(prefit, contrasts) %>% eBayes(trend = T, robust = T)
```


```{r eval=F}
limma_res <- purrr::map_dfr(
  colnames(contrasts),
  ~ topTable(fit, n=Inf, sort.by="none", coef=.x, confint = F  ) %>%
  rownames_to_column("GeneIDSite") %>%
  mutate(contrast=.x)
)
.file <- file.path(ROOTPATH, "MSPC000936_limma_res.tsv")
limma_res %>% write_tsv(.file)
# %>%
  #rownames_to_column("GeneIDSite") %>%
  #tibble() %>%
  #arrange(contrast, P.Value) #move from here to plot topN

#limma_res %>% head(20)
```


```{r, label=phos.geneset.heatmap, fig.width=12, fig.height=7}

.smarcs_h <- phos_gct_zscore@rdesc %>% filter(str_detect(GeneSymbol, "SMARC"))
.smarcs_m <- phos_gct_zscore@rdesc %>% filter(str_detect(GeneSymbol, "Smarc"))
.smarcs <- c(.smarcs_h$id,
             .smarcs_m$id
             )
pos_sel <- c("Extract_Treatment", "Extract_Genotype")
cut_by <- "Extract_Treatment"
cut_by <- "Extract_Genotype"
subset_gct(phos_gct_zscore, rid=.smarcs) %>% make_heatmap(show_row_names=T, pos_sel = pos_sel, cut_by = cut_by )
```



## top50 { .tabset }

### ints6 vs wt

```{r, fig.width=10, fig.height=9, eval=F}
.c <- "genotypeIntS6 - genotypeWT"
.top50 <- limma_res %>%
  filter(contrast == .c) %>%
  arrange(P.Value) %>%
  head(75)


subset_gct(phos_zscore, rid=.top50$GeneIDSite) %>% make_heatmap(show_row_names=T)

```

### ints12 vs wt

```{r, fig.width=10, fig.height=9, eval=F}
.c <- "genotypeIntS12 - genotypeWT"
.top50 <- limma_res %>%
  filter(contrast == .c) %>%
  arrange(P.Value) %>%
  head(75)


subset_gct(phos_zscore, rid=.top50$GeneIDSite) %>% make_heatmap(show_row_names=T)

```





```{r, fig.width=10, fig.height=9, eval=F}
.c <- "genotypeIntS6 - genotypeIntS12"
.top50 <- limma_res %>%
  filter(contrast == .c) %>%
  arrange(P.Value) %>%
  head(75)


subset_gct(phos_zscore, rid=.top50$GeneIDSite) %>% make_heatmap(show_row_names=T)

```


```{r, label=phos.params}
( subparams )

```

